---
title: "Tutorial RMarkdown"
output:
  html_document: default
  pdf_document: default
---

## R-Ladies-Lausanne

R Markdown is based on Markdown, "an easy-to-read and easy-to-write plain text format, optionally convert it to structurally valid XHTML (or HTML)" (John Gruber & Aaron Schwartz). In RStudio, an integrated development environment for R provides a C++ wrapper function for a markdown variant called sundown. _[source: Wikipedia]_

You can download the Cheat Sheet I distribute [here](https://www.rstudio.com/wp-content/uploads/2016/03/rmarkdown-cheatsheet-2.0.pdf)

__The structure of my R Markdown files:__

+ 1. text introduction to analysis
+ 2. initialization/setup
+ 3. code chunks and output

1. My _R Markdown_ files usually begin with some information in __Markdown__ format (as above) which describes the analysis. Useful information can be found at [rmarkdown](https://rmarkdown.rstudio.com)

2. Then, in the first 'chunk' (block of R code), I usually set the path to the folder in which I work, load necessary packages, define functions, etc. In the settings in the first line of the chunk I suppress showing the input (echo=FALSE) and the output (results='hide') in the final document. To avoid wordy warnings in the output, I suppress messages and warnings (warning=FALSE, message=FALSE). However, they will be displayes during creating 'knitting' the document. If I work on a simple dataset that doesn't need any processing, I also load it in this chunk. 

3. After this the actual code with the interesting output comes

## Initialization

```{r initialization, warning=FALSE, message=FALSE, echo=FALSE, results='hide'}
#setwd("/Users/claudiakasper/Desktop/")
library(lmerTest)
library(lme4)
library(ggplot2)
library(knitr)
data("iris")

simple_function <- function(a, b) { #this is a simple function
  c <- a + b
  print(c)
}

std.dev <- function(datavec) {
  std <- sqrt(var(datavec))
  print(std)
}
```

_Neither input nor any output of this chunk is displayed in the final document_

## Model Output

```{r summary}
summary(lmer(Petal.Length ~ Sepal.Length + (1|Species), data = iris))
```

If the whole output is not interesting, or if you want to display some additional statistics in the text, you can print them as in-line code: The mean sepal width across iris species is `r round(mean(iris$Sepal.Width), 2)` mm and the standard deviation 
is `r round(std.dev(iris$Sepal.Width), 2)`.

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
p <- ggplot(data = iris, aes(x=Sepal.Length, y=Petal.Length)) + # , colour = Species
            geom_point(colour = "darkgray") + geom_smooth(method = lm, colour = "black") 
p + theme_light(base_size = 11, base_family = "Helvetica") + facet_grid(cols = vars(Species))
```

## Tables

```{r table, echo=FALSE}
kable(iris[1:10,1:4], caption = "This is how a table can look like")
```

## Extras:

+ collapsing/navigating

+ figure size, etc

+ RStudio v1.2.1047 for presentations as PDF

+ LaTeX mathematical formulas

$R = h^2 * S$

+ [web page](https://bookdown.org/yihui/blogdown/), I am trying (but failing so far) with [hugo academic](https://themes.gohugo.io/academic/) 

+ fun stuff, e.g. [fontawesome](https://github.com/rstudio/fontawesome), [icon](https://github.com/ropenscilabs/icon), [anicon] (https://github.com/emitanaka/anicon)


```{r warning=FALSE, message=FALSE}
#install.packages("devtools")
#install_github("rstudio/fontawesome")
#install_github("ropenscilabs/icon")
#install_github("emitanaka/anicon")


library(fontawesome)
library(anicon)
library(icon)

anicon::faa("rocket", size = 5, animate = "spin")
anicon::faa("rocket", size = 5, rotate = 90)

icon::fa("hourglass-start")
icon::fa("hourglass-half")
icon::fa("hourglass-end")

ai_iconList # to see which icons are available
fa_iconList
```

also see http://jpswalsh.github.io/academicons/ where you can even request a new icon by creating an issue

local({r <- getOption("repos")
      r["CRAN"] <- "https://stat.ethz.ch/CRAN/"#       options(repos=r)})
